---
title: "Soil Moisture Probe Calibration"
author: "Long"
date: "2023-10-20"
output:
  html_document:
    theme: cerulean
---

This R workflow provides a demonstration of soil water calibration using two version of the SLGA data (Old and New). The calibration is performed with three distinct Pedo-transfer Functions (PTFs). We have employed a set of PTFs originally developed at the USYD and referred to as 'USYD JOSE' (Padarian, 2014), and have also conducted a comparative analysis with PTFs developed by the CSIRO (Somarathna and Searle, 2022). In addition, the soil water layers from Water Product 7 (WP7), where we estimated and mapped the drained upper limit (DUL) and critical lower limit (CLL) at each sampling location, with a spatial resolution of 90 meters were also introduced. The soil attributes employed for calibration purposes encompass soil composition, including sand, silt, and clay content, as well as bulk density and organic carbon. In this demonstration, 'PTF 1' signifies the CSIRO PTF, 'PTF 2' corresponds to the USYD JOSE PTF, and 'PTF 3' is associated with the CSIRO 90-meter product extraction, Sensor range is classified as Sensor 1 (0-30cm); Sensor 2 (30-60cm); Sensor 3 (60-100cm).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Package Loading
library(dplyr)
library(forecast)
library(here)
library(ggplot2)
library(leaflet)
library(raster)
library(readr)
library(rgdal)
library(rgeos)
library(sf)
library(sp)
library(tidyverse)
library(tidyr)
library(tseries)
source("C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/Accuracy.R")
```

##Part 1. OzNet API Download

```{r}
#Access data generation from the soil probe analysis and calibration
get_available_data <- function(){
  readRDS(file = here::here("C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/available_data.rds"))
}

check_SiteID_available <- function(SiteID) {
  available_data <- readRDS(file = here::here("C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/available_data.rds"))
  SiteID %in% available_data$SiteID
}
#' read_data
#'
#' @param siteID Name of the soil moisture probe, e.g. "OzNet_y1"
#'
#' @return Returns a dataframe of moisture and calibrated values(i.e rescaled column) for each depth
#' @export
#'
#' @examples
#' probe_data <- read_data('OzNet_y1')
read_data <- function(site) {
  if (check_SiteID_available(site) == TRUE) {
  all_data <- readRDS(here::here("C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/results.rds"))
  subset <- all_data %>% filter(SiteID == site) %>% pull(data)
  subset <- subset[[1]]
  return(subset)
  }
  else {
   # print("Error : Site Not available -Option must be within available_data.rds")
    return(stop(paste0("Error : Site",SiteID," Not available. Option must be within available_data.rds")))
  }

}
#' get_csvdata
#'
#' @param siteID Name of the soil moisture probe, e.g. "OzNet_y1"
#'
#' @return Returns a csv saved in the current working directory, whose name starts with siteID
#' @export
#'
#' @examples
#' get_csvdata('Cosmoz_10')
get_csvdata <- function(SiteID) {
  read_data(SiteID) %>% write_csv(file = here::here(paste0(SiteID,"_data.csv")))
}
#' get_highlevel_probedata
#'
#' @return Returns a dataframe of ProbeIDs that have are available, and high level summary data
#' @export
#'
#' @examples
#' available_data <- get_highlevel_probedata()
get_highlevel_probedata <- function(){
  available_data <- readRDS(file = here::here("C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/available_data.rds"))
  return(available_data)
}

available_data <- get_highlevel_probedata()

probe_data <- read_data('OzNet_y1') #Repeat for all avaiable probes from the network

write.csv(probe_data, "C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/OzNet/Yanco/OzNet_y1.csv", row.names = FALSE)

```

##Part 2. Probe Network Mapping (OzNet_Yanco as an example)

```{r}
#Load shapefile
Yanco <- st_read("C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/OzNet/Yanco/Shapefile")

#Load the CSV file with sampling sites
Probe_Y <- read.csv("C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/OzNet/Yanco/Probe_Yanco.csv")

#CRS adjustment
Probe_Y$Lon <- as.numeric(as.character(Probe_Y$Longitude))
Probe_Y$Lat <- as.numeric(as.character(Probe_Y$Latitude))
OzNet_Probes <- SpatialPoints(Probe_Y[, c("Lon", "Lat")])
proj4string(OzNet_Probes) <- CRS("+proj=longlat +datum=WGS84")
OzNet_Probes <- spTransform(OzNet_Probes,CRS("+proj=longlat +datum=WGS84"))
OzNet_Probes <- as.data.frame(OzNet_Probes, long = Longitude, lat = Latitude)

#Create a leaflet map
map <- leaflet() %>%
  addProviderTiles("Esri.WorldImagery", group = "Google Earth-like Basemap") %>%
  addPolygons(data = Yanco, color = "blue", fillOpacity = 0.4, group = "Yanco Shapefile")

#Add sampling sites to the map
map <- map %>%
  addCircleMarkers(
    data = Probe_Y,
    lat = ~Latitude,  # Replace with the actual column name in your CSV file
    lng = ~Longitude, # Replace with the actual column name in your CSV file
    color = "red",    # Customize marker color
    radius = 3        # Customize marker size
  )

#Display the map
map
```

##Part 3. PTFs Comparison & PAWC Calculation
#Input SLGA data and Extraction

```{r}
#SLGA Old extraction
soil.list1 <- list.files(path="C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/SLGAold", pattern =".tif", full.names=TRUE)
soil.stack1 <- raster::stack(soil.list1)

soilSLGA1 <- raster::extract(soil.stack1, OzNet_Probes)
soilSLGA1 <- as.data.frame(soilSLGA1)

names(soilSLGA1)[1:30]<-c( "BDW_005","BDW_015","BDW_030","BDW_060","BDW_100","BDW_200",
                          
                          "CLY_005","CLY_015","CLY_030","CLY_060","CLY_100","CLY_200",
                          
                          "SLT_005","SLT_015","SLT_030","SLT_060","SLT_100","SLT_200",
                          
                          "SND_005","SND_015","SND_030","SND_060","SND_100","SND_200",
                          
                          "SOC_005","SOC_015","SOC_030","SOC_060","SOC_100","SOC_200")

str(soilSLGA1)

#SLGA new extraction
soil.list2 <- list.files(path="C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/SLGAnew", pattern =".tif", full.names=TRUE)
soil.stack2 <- raster::stack(soil.list2)

soilSLGA2 <- raster::extract(soil.stack2, OzNet_Probes)
soilSLGA2 <- as.data.frame(soilSLGA2)

names(soilSLGA2)[1:30]<-c( "BDW_005","BDW_015","BDW_030","BDW_060","BDW_100","BDW_200",
                          
                          "CLY_005","CLY_015","CLY_030","CLY_060","CLY_100","CLY_200",
                          
                          "SLT_005","SLT_015","SLT_030","SLT_060","SLT_100","SLT_200",
                          
                          "SND_005","SND_015","SND_030","SND_060","SND_100","SND_200",
                          
                          "SOC_005","SOC_015","SOC_030","SOC_060","SOC_100","SOC_200")

str(soilSLGA2)
```

#Read Soil Moisture Data

```{r}
OzNet_Y <- read.csv("C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/OzNet/Yanco/Yanco_SM.csv")
str(OzNet_Y)

#Back transform reported SM to Normalized Raw Counts
Rawcounts <- function(Theta,CoefA=0.232,ExpB=0.41,ConC=-0.021){
  RC <- CoefA*Theta^(ExpB)+ConC
  return(RC)
}
OzNet_Y$NRC <- Rawcounts(Theta = OzNet_Y$SM,CoefA = 0.1957,ExpB = 0.404,ConC = 0.02852)
OzNet_Y$defaultCal <- (((Rawcounts(Theta = OzNet_Y$SM,CoefA = 0.1957,ExpB = 0.404,ConC = 0.02852)-(-0.021))/0.232)^(1/0.41))/100

#Identify Maxima and minima 
SensorRange <- OzNet_Y %>% group_by(Sensor,Site_ID) %>%
  summarise(max=max(defaultCal),min=min(defaultCal))

OzNet_Y <- left_join(OzNet_Y,SensorRange)
```

#SM Stretch PTF 1 - CSIRO PTF

```{r}
soil_data <- read.csv("C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/OzNet/Yanco/SoilSLGA_Y.csv")

#Old SLGA Data
StretchPars1_O <- function(soil_data){
  theta_dul1_O = ((48.98 - 21.86 * soil_data$BDW_1 + 0.36 * soil_data$CLY_1 - 0.06 * soil_data$SND_1 - 0.19 * sqrt(soil_data$SLT_1) + 2.90 * sqrt(soil_data$SOC_1))/100)
  
  theta_pwp1_O<- ((17.40 - 10.05 * soil_data$BDW_1 + 0.34 * soil_data$CLY_1 - 0.02 * soil_data$SND_1 + 0.18 * soil_data$SLT_1)/100)
  
  return(cbind(as.character(soil_data$Site_ID),soil_data$Sensor,theta_dul1_O,theta_pwp1_O))
}

SLGA_Bucketbounds1_O <- as.data.frame(StretchPars1_O(soil_data))
names(SLGA_Bucketbounds1_O) <- c("Site_ID","Sensor","SLGA_DUL1_O","SLGA_PWP1_O")
str(SLGA_Bucketbounds1_O)

#New SLGA Data
StretchPars1_N <- function(soil_data){
  theta_dul1_N = ((48.98 - 21.86 * soil_data$BDW_2 + 0.36 * soil_data$CLY_2 - 0.06 * soil_data$SND_2 - 0.19 * sqrt(soil_data$SLT_2) + 2.90 * sqrt(soil_data$SOC_2))/100)
  
  theta_pwp1_N<- ((17.40 - 10.05 * soil_data$BDW_2 + 0.34 * soil_data$CLY_2 - 0.02 * soil_data$SND_2 + 0.18 * soil_data$SLT_2)/100)
  
  return(cbind(as.character(soil_data$Site_ID),soil_data$Sensor,theta_dul1_N,theta_pwp1_N))
}

SLGA_Bucketbounds1_N <- as.data.frame(StretchPars1_N(soil_data))
names(SLGA_Bucketbounds1_N) <- c("Site_ID","Sensor","SLGA_DUL1_N","SLGA_PWP1_N")
str(SLGA_Bucketbounds1_N)
```

#SM Stretch PTF 2 - USYD JOSE PTF

```{r}
#Old SLGA Data
StretchPars2_O <- function(soil_data){
  theta_dul2_O = 0.4795 - 3.873 * 10^-5 * soil_data$SND_1 ^2 - 6.701 * 10^-7 * soil_data$CLY_1 ^2 * soil_data$SND_1
  
  theta_pwp2_O<- -0.1554 - 0.7221 * tanh(0.5 * (-0.9705 - 0.8529 * soil_data$BDW_1 - 0.00827 *
                                               soil_data$CLY_1 + 0.01994 * soil_data$SND_1))  + 0.1325 * tanh(0.5 * (3.71 - 3.19 * soil_data$BDW_1 
                                                                                                                   + 0.01205 * soil_data$CLY_1 + 0.01617 * soil_data$SND_1)) + 0.1720 * tanh(0.5 * (-3.94 - 0.5067 * soil_data$BDW_1
                                                                                                                                                                                                  + 0.02158 * soil_data$CLY_1 + 0.04978 * soil_data$SND_1))
  return(cbind(as.character(soil_data$Site_ID),soil_data$Sensor,theta_dul2_O,theta_pwp2_O))
}

SLGA_Bucketbounds2_O <- as.data.frame(StretchPars2_O(soil_data))
names(SLGA_Bucketbounds2_O) <- c("Site_ID","Sensor","SLGA_DUL2_O","SLGA_PWP2_O")
str(SLGA_Bucketbounds2_O)

#New SLGA Data
StretchPars2_N <- function(soil_data){
  theta_dul2_N = 0.4795 - 3.873 * 10^-5 * soil_data$SND_2 ^2 - 6.701 * 10^-7 * soil_data$CLY_2 ^2 * soil_data$SND_2
  
  theta_pwp2_N<- -0.1554 - 0.7221 * tanh(0.5 * (-0.9705 - 0.8529 * soil_data$BDW_2 - 0.00827 *
                                               soil_data$CLY_2 + 0.01994 * soil_data$SND_2))  + 0.1325 * tanh(0.5 * (3.71 - 3.19 * soil_data$BDW_2 
                                                                                                                   + 0.01205 * soil_data$CLY_2 + 0.01617 * soil_data$SND_2)) + 0.1720 * tanh(0.5 * (-3.94 - 0.5067 * soil_data$BDW_2
                                                                                                                                                                                                  + 0.02158 * soil_data$CLY_2 + 0.04978 * soil_data$SND_2))
  return(cbind(as.character(soil_data$Site_ID),soil_data$Sensor,theta_dul2_N,theta_pwp2_N))
}

SLGA_Bucketbounds2_N <- as.data.frame(StretchPars2_N(soil_data))
names(SLGA_Bucketbounds2_N) <- c("Site_ID","Sensor","SLGA_DUL2_N","SLGA_PWP2_N")
str(SLGA_Bucketbounds2_N)
```

#SM Stretch PTF 3 - CSIRO 90M product

```{r}
StretchPars3 <- list.files(path="C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/CSIRO90m", pattern =".tif", full.names=TRUE)
StretchPars3 <- raster::stack(soil.list2)

soilCSIRO <- raster::extract(StretchPars3, OzNet_Probes)
soilCSIRO <- as.data.frame(soilCSIRO)

#Adapt to DUL&CLL layer names
names(soilCSIRO)[1:12]<-c( "CLL_005","CLL_015","CLL_030","CLL_060","CLL_100","CLL_200",
                           
                           "DUL_005","DUL_015","DUL_030","DUL_060","DUL_100","DUL_200")

DUL_005 <- (soilCSIRO$DUL_005)/100
DUL_015 <- (soilCSIRO$DUL_015)/100
DUL_030 <- (soilCSIRO$DUL_030)/100
DUL_060 <- (soilCSIRO$DUL_060)/100
DUL_100 <- (soilCSIRO$DUL_100)/100
DUL_200 <- (soilCSIRO$DUL_200)/100
PWP_005 <- (soilCSIRO$CLL_005)/100
PWP_015 <- (soilCSIRO$CLL_015)/100
PWP_030 <- (soilCSIRO$CLL_030)/100
PWP_060 <- (soilCSIRO$CLL_060)/100
PWP_100 <- (soilCSIRO$CLL_100)/100
PWP_200 <- (soilCSIRO$CLL_200)/100

SLGA_Bucketbounds3 <- cbind(DUL_005, DUL_015, DUL_030, DUL_060, DUL_100, DUL_200,
                            PWP_005, PWP_015, PWP_030, PWP_060, PWP_100, PWP_200)

SLGA_Bucketbounds3 <- as.data.frame(SLGA_Bucketbounds3)
str(SLGA_Bucketbounds3)
```

## Read in probe timeseries

```{r}
LinNormDF <- read.csv("C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/OZnet/Yanco/OzNet_Y.csv")
LinNormDF$SM <- LinNormDF$SM/100
str(LinNormDF)

#Create LN function
linnorm <- function(ProbeMax,ProbeMin,ProbeTheta,DUL,PWP){
  Theta <- ((ProbeTheta-ProbeMin)/(ProbeMax-ProbeMin))*(DUL-PWP)+PWP
  return(Theta)
}

#Linear normalisation
LinNormDF <- na.omit(LinNormDF)
LinNormDF$Date <- as.Date(LinNormDF$Date)
LinNormDF$Sensor <- as.factor(LinNormDF$Sensor)
LinNormDF$Site_ID <- as.numeric(LinNormDF$Site_ID)

LinNormDF$SLGANorm1_O <- linnorm(ProbeMax = LinNormDF$Max,
                               ProbeMin = LinNormDF$Min,
                               ProbeTheta = LinNormDF$defaultCal,
                               DUL=LinNormDF$DUL1_O,
                               PWP=LinNormDF$PWP1_O)

LinNormDF$SLGANorm1_N <- linnorm(ProbeMax = LinNormDF$Max,
                               ProbeMin = LinNormDF$Min,
                               ProbeTheta = LinNormDF$defaultCal,
                               DUL=LinNormDF$DUL1_N,
                               PWP=LinNormDF$PWP1_N)

LinNormDF$SLGANorm2_O <- linnorm(ProbeMax = LinNormDF$Max,
                               ProbeMin = LinNormDF$Min,
                               ProbeTheta = LinNormDF$defaultCal,
                               DUL=LinNormDF$DUL2_O,
                               PWP=LinNormDF$PWP2_O)

LinNormDF$SLGANorm2_N <- linnorm(ProbeMax = LinNormDF$Max,
                               ProbeMin = LinNormDF$Min,
                               ProbeTheta = LinNormDF$defaultCal,
                               DUL=LinNormDF$DUL2_N,
                               PWP=LinNormDF$PWP2_N)

LinNormDF$SLGANorm3 <- linnorm(ProbeMax = LinNormDF$Max,
                               ProbeMin = LinNormDF$Min,
                               ProbeTheta = LinNormDF$defaultCal,
                               DUL=LinNormDF$DUL_3,
                               PWP=LinNormDF$PWP_3)

#Time-series Plot
ggplot(data=LinNormDF,aes(x=Date,y=defaultCal,colour=Sensor))+
  geom_path()+
  facet_wrap(vars(Site_ID))

SLGA_LNTS <- ggplot(data=LinNormDF,aes(x=Date,y=SLGANorm1_O,colour=Sensor))+
  geom_path()+
  facet_wrap(vars(Site_ID))+
  geom_point(data=LinNormDF,aes(x=Date,y=SM,colour=Sensor))+
  labs(x="Date",
       y= bquote("SLGA Linear Normalisation"*theta[V]),
       title= "SLGA Linear Normalisation Timeseries")+
  scale_x_date(date_breaks = "2 years", date_labels = "%Y")+
  theme(axis.text.x = element_text(size=10)) #Repeat for SLGA2 and 3
print(SLGA_LNTS)

```

##LN Plot

```{r}
LinNormDF$Sensor <- as.numeric(as.character(LinNormDF$Sensor))
LNormOBSPRED <- LinNormDF
LNormOBSPRED <- na.omit(LNormOBSPRED)
LNormOBSPRED$Site_ID <- as.factor(as.numeric(as.factor(LNormOBSPRED$Site_ID)))
str(LNormOBSPRED)

write.csv(LNormOBSPRED, "C:/USYD_SoilMoisture_Calibration/Soil Prob calibration project/USYD_Calibration_2023/OZnet/Yanco/OzNet_Yanco_calibrated.csv", row.names = FALSE)

#Correlation Test
#PTF1 SLGA Old
SLGANormFit1 <- LNormOBSPRED %>% group_by(Site_ID) %>% summarise(ME=accuracy(SM,SLGANorm1_O)[1],
                                                                RMSE=accuracy(SM,SLGANorm1_O)[2],
                                                                R2=accuracy(SM,SLGANorm1_O)[3],
                                                                LCCC=accuracy(SM,SLGANorm1_O)[4])

SLGAnormGoof1 <- accuracy(actual = LNormOBSPRED$SM,predict = LNormOBSPRED$SLGANorm1_O)

#PTF1 SLGA New
SLGANormFit2 <- LNormOBSPRED %>% group_by(Site_ID) %>% summarise(ME=accuracy(SM,SLGANorm1_N)[1],
                                                                RMSE=accuracy(SM,SLGANorm1_N)[2],
                                                                R2=accuracy(SM,SLGANorm1_N)[3],
                                                                LCCC=accuracy(SM,SLGANorm1_N)[4])

SLGAnormGoof2 <- accuracy(actual = LNormOBSPRED$SM,predict = LNormOBSPRED$SLGANorm1_N)

#PTF2 SLGA Old
SLGANormFit3 <- LNormOBSPRED %>% group_by(Site_ID) %>% summarise(ME=accuracy(SM,SLGANorm2_O)[1],
                                                                RMSE=accuracy(SM,SLGANorm2_O)[2],
                                                                R2=accuracy(SM,SLGANorm2_O)[3],
                                                                LCCC=accuracy(SM,SLGANorm2_O)[4])

SLGAnormGoof3 <- accuracy(actual = LNormOBSPRED$SM,predict = LNormOBSPRED$SLGANorm2_O)

##PTF2 SLGA New
SLGANormFit4 <- LNormOBSPRED %>% group_by(Site_ID) %>% summarise(ME=accuracy(SM,SLGANorm2_N)[1],
                                                                RMSE=accuracy(SM,SLGANorm2_N)[2],
                                                                R2=accuracy(SM,SLGANorm2_N)[3],
                                                                LCCC=accuracy(SM,SLGANorm2_N)[4])

SLGAnormGoof4 <- accuracy(actual = LNormOBSPRED$SM,predict = LNormOBSPRED$SLGANorm2_N)

#PTF3 SLGA
SLGANormFit5 <- LNormOBSPRED %>% group_by(Site_ID) %>% summarise(ME=accuracy(SM,SLGANorm3)[1],
                                                                RMSE=accuracy(SM,SLGANorm3)[2],
                                                                R2=accuracy(SM,SLGANorm3)[3],
                                                                LCCC=accuracy(SM,SLGANorm3)[4])

SLGAnormGoof5 <- accuracy(actual = LNormOBSPRED$SM,predict = LNormOBSPRED$SLGANorm3)

#Plotting
SLGALN_Goof_plot <- ggplot(data=LNormOBSPRED,aes(x=SM,y=SLGANorm1_O,colour=Site_ID))+
  geom_point()+
  geom_abline()+xlim(0,0.5)+ylim(0,0.5)+geom_text(colour="black",aes(x=-Inf, y=Inf, hjust=0, vjust=1,
                                                                     label=paste(" R2 =", formatC(SLGAnormGoof1[3],digits = 3,format="f"), "\n",
                                                                                 "RMSE =", format(SLGAnormGoof1[2],digits = 3,format="f"), "\n")))+
  labs(x= bquote("Observed Moisture "*theta[V]),
       y= bquote("SLGA Data Linear Normalised Moisture "*theta[V]),
       title= "SLGA Data Linear Normalised Moisture All Sites",
       legend="Site No")

print(SLGALN_Goof_plot) #Repeat for SLGA2 and 3
